
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, comment = FALSE, error = FALSE, message = FALSE)
```

```{r cars}
library(lubridate)
library(tidyverse)
library(MASS)
library(broom)
library(mgcv)
library(RcppRoll)
library(dplyr)
library(skimr)
library(magrittr)
library(zoo)
```

#### Import dataset
```{r pressure, echo=FALSE}
vic_weather <- read.csv("vic_weather.csv")
```

#### Preprocess the car_accident_victoria dataset 
```{r}
# Clean up the column names
cav_data_link <- "/Users/merry/Documents/Deakin - Master of Data Science/T2 2022/SIT741/Assessments/A3/car_accidents_victoria.csv"
top_row <- read_csv(cav_data_link, col_names = FALSE, n_max = 1)
second_row <- read_csv(cav_data_link, n_max = 1)

column_names <- second_row %>%
  unlist(., use.names = FALSE) %>%
  make.unique(., sep = "__") # double underscore

column_names[2:5] <- str_c(column_names[2:5], '0', sep='__')

daily_accidents <-
  read_csv(cav_data_link, skip = 2, col_names = column_names)
```

```{r eval=TRUE}
# Get the regions
(
regions <- top_row %>% 
  unlist(., use.names=FALSE) %>% 
  na.omit()
)
```

#### Transform dataset into tidy data
```{r}
daily_accidents <- daily_accidents %>% 
  gather(key = index, 
         value = value, 
         -DATE)
```

```{r}
# Put data from the same measurement source in a column and put each observation in a row
daily_accidents <- daily_accidents %>% 
  separate(index, 
           into = c("index", 
                    "region_id"), 
           sep="__", 
           remove=TRUE) %>% 
  mutate(region_id = 
           as.numeric(region_id) + 1) %>% 
  mutate(region_name = 
           regions[region_id]) %>% 
  dplyr::select(DATE, 
         region_name, 
         index, 
         value)
```

```{r}
# Replacing N/a values with the mode 
getmode <- function(v) {
   v <- na.omit(v)
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}


daily_accidents <- daily_accidents %>% 
  mutate(value = 
           as.numeric(value))

daily_accidents <- daily_accidents %>% 
  mutate(value = replace_na(value, getmode(value)))


daily_accidents <- daily_accidents %>% 
  spread(index, 
         value)
```

```{r}
daily_accidents <- daily_accidents %>% 
  mutate(DATE = dmy(DATE))
```

#### Select a region for the following analysis
```{r}
selected_region <- "METROPOLITAN NORTH WEST REGION"

one_region <- daily_accidents %>%
  filter(region_name == selected_region) %>% 
  dplyr::select(-region_name) %>% 
  arrange(DATE)

one_region %>% 
  summarise(total_accidents = sum(SERIOUS)+sum(FATAL)+sum(NOINJURY)+sum(OTHER), 
            serious_accidents = sum(SERIOUS))

one_region <- one_region %>%
  mutate(TOTAL_ACCIDENTS = FATAL + SERIOUS + NOINJURY + OTHER)

one_region
```

```{r}
one_region <- transform(one_region, 
                        ndate = as.numeric(DATE),
                        nyear  = as.numeric(format(DATE, '%Y')),
                        nmonth = as.numeric(format(DATE, '%m')),
                        day = as.numeric(format(DATE, '%j')))
str(one_region)
```

#### Fit a Poisson distribution and a negative binomial distribution on the sample of total accidents
```{r}
region_sample <- sample_n(one_region, 200)

fit_P <- fitdist(region_sample$TOTAL_ACCIDENTS, "pois")

fit_NB <- fitdist(region_sample$TOTAL_ACCIDENTS, "nbinom")

denscomp(list(fit_P, fit_NB))

cdfcomp(list(fit_P, fit_NB))

print(region_sample$TOTAL_ACCIDENTS)

```

#### Compare the log-likelihood of two fitted distributions.
```{r}
one_region %>% 
  ggplot(aes(sample = TOTAL_ACCIDENTS)) +
  stat_qq(distribution = stats::qnbinom, dparams = fit_NB$estimate) +
  stat_qq_line(distribution = stats::qnbinom, dparams = fit_NB$estimate)
```
```{r}
fit_P$loglik
fit_NB$loglik
```

#### Fit a linear model for road traffic accident using date as the predictor variable
```{r}
# We can first plot the total accidents number against the date of a year (day)
one_region %>% 
  ggplot(aes(x = day, 
             y = TOTAL_ACCIDENTS)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
# Then, fit the linear model and assess the model fit
model_lm <- lm(TOTAL_ACCIDENTS ~ day,
               data = one_region)
model_lm
```

```{r}
# Obtain model fitting summary
model_lm %>%
  summary
```

```{r}
model_lm %>%
  glance
```

#### Plot the fitted values and the residuals
```{r}
results <- model_lm %>%
  augment
```

```{r}
results %>% 
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_smooth(colour = 'firebrick') 
```

#### Fit a generalised additive model (GAM)
```{r}
model_gam <- gam(TOTAL_ACCIDENTS ~ s(day), data = one_region, method = "REML")
model_gam
```

```{r}
summary(model_gam)
```

```{r}
# Plot the partial effective scatter plot
plot(model_gam,seWithMean = TRUE, residuals = TRUE, shift = coef(model_gam)[1])
```

```{r}
par(mfrow = c(2, 2))
gam.check(model_gam)
```

#### Augment the model to incorporate the weekly variations
#### Investigate the weekly pattern 
```{r}
week_mnw <- one_region %>%
  mutate(WEEK = wday(DATE, label = TRUE))

week_mnw %>%
  ggplot(aes(x = WEEK,
             y = TOTAL_ACCIDENTS)) +
  geom_violin() +
  geom_boxplot(width = 0.2)
```

```{r}
week_mnw %>%
  ggplot(aes(x = DATE, y = TOTAL_ACCIDENTS, colour = factor(WEEK))) + 
  geom_line()
```

#### Fit the new model with weekly variations
```{r}
model_gam_week <- gam(TOTAL_ACCIDENTS ~ s(ndate, by = WEEK) + WEEK, data = week_mnw)
model_gam_week
```

#### Use plots to see the fitted smooth functions
```{r}
par(mfrow = c(2, 3))
plot(model_gam_week, shade = TRUE, shade.col = "lightblue")
```

#### Evaluate the model after incorporating weekly variations
```{r}
summary(model_gam_week)
```

```{r}
par(mfrow = c(2, 2))
gam.check(model_gam_week)
```

#### Compare the models using AIC
```{r}
AIC(model_lm, model_gam, model_gam_week)
```

```{r}
print("Linear Model:")
coef(model_lm)
print("GAM Model:")
coef(model_gam)
print("GAM Model with weekly variations:")
coef(model_gam_week)
```

```{r}
anova(model_gam, model_gam_week)
```

### Heatwaves, precipitation and road traffic accidents

#### Use the NOAA data to calculate the dialy EHF and plot the daily values
```{r}
# Calculate the average temperature of the day using min-max values
vic_ehf <- vic_weather %>%
  mutate(TMEAN = (TMAX + TMIN)/2)
```

```{r}
# Add the columns of three-day-average DMT and average DMT over the recent past (30 days)
mnw_ehf <- vic_ehf %>%
  mutate(t_3d_after = roll_meanr(lead(TMEAN), 3)) %>%
  mutate(t_30d_before = roll_meanr(lag(TMEAN), 30))
```

```{r}
# Calculate T95 for each year and add the value to the dataset
mnw_ehf <- transform(mnw_ehf, nyear = as.numeric(format(DATE, '%Y')))

t95 <- mnw_ehf %>%
  group_by(nyear) %>%
  summarise(t_95 = quantile(TMEAN, probs = 0.95, na.rm = TRUE))
  
mnw_ehf <- merge(mnw_ehf, t95, by = "nyear")
```

```{r}
# Calculate the EHI
mnw_ehf <- mnw_ehf %>%
  mutate(EHIsig = t_3d_after - t_95) %>%
  mutate(EHIaccl = t_3d_after - t_30d_before) %>%
  mutate(EHF = EHIsig * pmax(1, EHIaccl))
```

```{r}
# Plot the daily EHI
mnw_ehf %>%
  ggplot(aes(x=DATE, 
             y=EHF)) + 
  geom_line() + 
  ylab("Excess Heat Factor (EHF)")
```

#### Use the EHF as an additional predictor to augment the models
```{r}
# Make a new dataframe for the GAM model fit
mnw_ehf <- mnw_ehf %>%
  mutate(WEEK = wday(DATE, label = TRUE))

mnw_ehf_cb <- merge(mnw_ehf, week_mnw, by = c("DATE", "nyear", "WEEK"))
```

```{r}
# Fit the GAM model with weekly variation and EHF predictors
model_gam_ehf <- gam(TOTAL_ACCIDENTS ~ s(ndate, by = WEEK) + WEEK + s(EHF), data = mnw_ehf_cb)
model_gam_ehf
```

```{r}
summary(model_gam_ehf)
```

```{r}
par(mfrow = c(2, 2))
gam.check(model_gam_ehf)
```
```{r}
# Plot the results
par(mfrow = c(2, 4))
plot(model_gam_ehf, shade = TRUE, shade.col = "lightblue")
```
```{r}
# Use the shift argument to shift the scale for an interpretable purpose. 
plot(model_gam_ehf, seWithMean = TRUE, shade = TRUE, shade.col = "lightblue", shift = coef(model_gam_ehf)[1])
```

```{r}
# Compare AIC scores of the all the models
AIC(model_lm, model_gam, model_gam_week, model_gam_ehf)
```

#### Use extra weather features for the mode
```{r}
# Fit the model with daily minimum and maximum temperature
model_gam_minmax<- gam(TOTAL_ACCIDENTS ~ s(ndate, by = WEEK) + WEEK + te(TMIN, TMAX), data = mnw_ehf_cb)
model_gam_minmax
```

```{r}
# Fit the model with daily precipitation
model_gam_prcp <- gam(TOTAL_ACCIDENTS ~ s(ndate, by = WEEK) + WEEK + s(EHF) + s(TAVG) + PRCP, data = mnw_ehf_cb)
model_gam_prcp
```

```{r}
summary(model_gam_minmax)
```

```{r}
summary(model_gam_prcp)
```

```{r}
par(mfrow = c(2, 2))
gam.check(model_gam_minmax)
```

```{r}
par(mfrow = c(2, 2))
gam.check(model_gam_prcp)
```
```{r}
# Compare all the AIC scores 
AIC(model_lm, model_gam, model_gam_week, model_gam_ehf,
    model_gam_minmax, model_gam_prcp)
```

#### The End ####


